---
title: "01_preparation_workingdata_BnSites"
author: "Bernd Panassiti"
date: "19 Mar 2015"
output: pdf_document
---

infection rates of combined sites need to be extracted by hand


# 64, 310

SELECT count(ill) FROM boisnoir_vines WHERE site = '64';
452

SELECT count(ill) FROM boisnoir_vines WHERE site = '64-310';
496

```{r}
set.seed(2013)
combinedSites<-cbind(1:496,sample(c("64", "310"), 496, replace=T, prob=c(0.5, 0.5)))
nrow(combinedSites[combinedSites[,2]=='64',])
# 225
nrow(combinedSites[combinedSites[,2]=='310',])
# 271
combinedSites64_310 <- combinedSites
```



# 70, 311

SELECT count(ill='t') FROM boisnoir_vines WHERE site = '70';
SELECT count(ill) FROM boisnoir_vines WHERE site = '70'AND ill='t';
287
1

SELECT count(ill) FROM boisnoir_vines WHERE site = '70-311';
2324

```{r}
selection <- bn.alt[bn.alt$site=='70-311',1:2]
set.seed(2013)
combinedSites<-cbind(selection,sample(c("70", "311"), 2324, replace=T, prob=c(0.5, 0.5)))
nrow(combinedSites[combinedSites[,3]=='70',])
# 1136
nrow(combinedSites[combinedSites[,3]=='70' & combinedSites[,2]==1,])
# 41

nrow(combinedSites[combinedSites[,3]=='311',])
# 1188
nrow(combinedSites[combinedSites[,3]=='311' & combinedSites[,2]==1,])
# 34

combinedSites70_311 <- combinedSites
```


# 27, 28

SELECT count(ill='t') FROM boisnoir_vines WHERE site = '27';
2132
SELECT count(ill) FROM boisnoir_vines WHERE site = '27'AND ill='t';
30

SELECT count(ill='t') FROM boisnoir_vines WHERE site = '28';
1049
SELECT count(ill) FROM boisnoir_vines WHERE site = '28'AND ill='t';
12

SELECT count(ill) FROM boisnoir_vines WHERE site = '27-28';
5

```{r}
set.seed(2013)
combinedSites27_28<-cbind(1:5,sample(c("27", "28"), 5, replace=T, prob=c(0.5, 0.5)))
```





# 36, 37

No infected vines.

SELECT count(ill='t') FROM boisnoir_vines WHERE site = '36';
2159

SELECT count(ill='t') FROM boisnoir_vines WHERE site = '37';
2007


SELECT count(ill) FROM boisnoir_vines WHERE site = '36-37';
114

```{r}
set.seed(2013)
combinedSites36_37<-cbind(1:114,sample(c("36", "37"), 114, replace=T, prob=c(0.5, 0.5)))
nrow(combinedSites36_37[combinedSites36_37[,2]=='36',])
# 58
nrow(combinedSites36_37[combinedSites36_37[,2]=='37',])
# 56
```



# 416, 417, 418

SELECT count(ill) FROM boisnoir_vines WHERE site = '416';
SELECT count(ill) FROM boisnoir_vines WHERE site = '416'AND ill='t';
633, 1

SELECT count(ill) FROM boisnoir_vines WHERE site = '417';
SELECT count(ill) FROM boisnoir_vines WHERE site = '417'AND ill='t';
688,6


SELECT count(ill) FROM boisnoir_vines WHERE site = '418';
SELECT count(ill) FROM boisnoir_vines WHERE site = '418'AND ill='t';
687,6


## 416-417
SELECT count(ill) FROM boisnoir_vines WHERE site = '416-417';
SELECT count(ill) FROM boisnoir_vines WHERE site = '416-417' AND ill='t';
238,1
```{r}
selection <- bn.alt[bn.alt$site=='416-417',1:2]
set.seed(2013)
combinedSites<-cbind(selection,sample(c("416", "417"), 238, replace=T, prob=c(0.5, 0.5)))
nrow(combinedSites[combinedSites[,3]=='416',])
# 117
nrow(combinedSites[combinedSites[,3]=='416' & combinedSites[,2]==1,])
# 1

nrow(combinedSites[combinedSites[,3]=='417',])
# 121
nrow(combinedSites[combinedSites[,3]=='417' & combinedSites[,2]==1,])
# 0

combinedSites416_417 <- combinedSites
```



## 416-418
SELECT count(ill) FROM boisnoir_vines WHERE site = '416-418';
SELECT count(ill) FROM boisnoir_vines WHERE site = '416-418' AND ill='t';
406,4

```{r}
selection <- bn.alt[bn.alt$site=='416-418',1:2]
set.seed(2013)
combinedSites<-cbind(selection,sample(c("416", "418"), 406, replace=T, prob=c(0.5, 0.5)))
nrow(combinedSites[combinedSites[,3]=='416',])
# 208
nrow(combinedSites[combinedSites[,3]=='416' & combinedSites[,2]==1,])
# 3

nrow(combinedSites[combinedSites[,3]=='418',])
# 198
nrow(combinedSites[combinedSites[,3]=='418' & combinedSites[,2]==1,])
# 1

combinedSites416_418 <- combinedSites
```





## 417-418
SELECT count(ill) FROM boisnoir_vines WHERE site = '417-418';
SELECT count(ill) FROM boisnoir_vines WHERE site = '417-418' AND ill='t';
349,5


```{r}
selection <- bn.alt[bn.alt$site=='417-418',1:2]
set.seed(2013)
combinedSites<-cbind(selection,sample(c("417", "418"), 349, replace=T, prob=c(0.5, 0.5)))
nrow(combinedSites[combinedSites[,3]=='417',])
# 177
nrow(combinedSites[combinedSites[,3]=='417' & combinedSites[,2]==1,])
# 3

nrow(combinedSites[combinedSites[,3]=='418',])
# 172
nrow(combinedSites[combinedSites[,3]=='418' & combinedSites[,2]==1,])
# 2

combinedSites417_418 <- combinedSites
```





## 416-417-418
SELECT count(ill) FROM boisnoir_vines WHERE site = '416-417-418';
SELECT count(ill) FROM boisnoir_vines WHERE site = '416-417-418' AND ill='t';
1175,8

```{r}
selection <- bn.alt[bn.alt$site=='416-417-418',1:2]
set.seed(2013)
combinedSites<-cbind(selection,sample(c("416", "417","418"), 1175, replace=T, prob=c(0.5, 0.5,0.5)))
nrow(combinedSites[combinedSites[,3]=='416',])
# 367
nrow(combinedSites[combinedSites[,3]=='416' & combinedSites[,2]==1,])
# 4

nrow(combinedSites[combinedSites[,3]=='417',])
# 419
nrow(combinedSites[combinedSites[,3]=='417' & combinedSites[,2]==1,])
# 2

nrow(combinedSites[combinedSites[,3]=='418',])
# 389
nrow(combinedSites[combinedSites[,3]=='418' & combinedSites[,2]==1,])
# 2
combinedSites416_417_418 <- combinedSites
```

416
633, 1
117,1
208,3
367,4
TOTAL: 1325, 9





417
633, 1
121,0
177,3
419,2
TOTAL: 1350, 6


418
687,6
198,1
172,2
389,2
TOTAL:1446,11




```{r}
save(
  combinedSites64_310,
  combinedSites70_311,
  combinedSites27_28,
  combinedSites36_37,
  combinedSites416_417,
  combinedSites416_418,
  combinedSites417_418,
  combinedSites416_417_418,
  file="data/hyal_altitude_BnSites.RData"
  )
```

